---
title: "Intro to RBioTools"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, echo = FALSE)
```

# Introduction to RBiotools

RBiotools is a package for COMPARATIVE MICROBIAL GENOMICS. RBiotools works best in the RStudio integrated development environment. RBiotools works with IDs from GenBank, the NIH genetic sequence data base, to perform analyses and create figures.

## Before installation, check the version of R that is being used
These instructions are prepared using:
*R version 4.1.1 (2021-08-10)


```{r verify.version, eval=TRUE, echo=FALSE}

sessionInfo()

```


## Install the RBiotools package and other required packages

1. install the dependencies
```{r install.dependencies, eval=FALSE, echo=TRUE}

install.packages(c("rlang","ape", "fmsb", "gplots", "grImport", "gridExtra", "msa", "pheatmap", "RCurl", "rentrez", "seqinr"))
install.packages(c("lattice","mgcv","nlme","survival"))

#Install msa and Biostrings package via BiocManager
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("msa")
BiocManager::install("Biostrings", force = TRUE)


install.packages("installr")
library(installr)


```


2. install the RBiotools package from zip source; and activate it in the workspace

```{r install.packages, eval=FALSE, echo=TRUE}

install.packages("C:/dc3/RBiotools_0.5.0.tar.gz", repos = NULL, type = "source")
library(RBiotools)

```

3. Setup the environment and working directory
Choose a directory / folder that will be used for some of the figures, including Heat Maps and Genome Atlases:

* Create this directory on your computer, if it does not already exist
* Make this directory the "Working Directory"

```{r working.directory, eval=FALSE, echo=FALSE}

#this line checks to see if the folder exists; if it does not then it creates the folder; if it does then it returns "Folder exists already"
ifelse(!dir.exists(".//r_repository//"), dir.create(".//r_repository//"), "Folder exists already")

setwd("C:/dc3/R_repository")

```

# Introduction to RBiotools

1. Choose a set of organisms
What is an organism?
* Organisms are specified with GenBank identifiers 
* Organisms NOT in Genbank can be added with the RBiotools addGenome function

Let's use a sample sets of organisms to start with:

In the following statement:
   "myGenomes" is an arbitrary variable name
   "c" is an R function to "combine" things into a list
   Comments are for the user's benefit and are NOT required
   Items in a list DO NOT have to be on separate lines


















